---
title: "tidyprint"
output: "html_vignette"
params:
  demo_metadata: true
vignette: >
  %\VignetteIndexEntry{posteriorHCA}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
knit: >
    (function(x, ...){
        proj_root <- rprojroot::find_package_root_file() |> normalizePath()
        rmarkdown::render(
            x,
            output_format = "html_document",
            params = list(
                demo_metadata = TRUE
            )
        )
        rmarkdown::render(
            x,
            output_file = "README.md",
            output_format = "github_document",
            output_dir = proj_root,
            knit_root_dir = proj_root,
            params = list(
                demo_metadata = FALSE
            )
        )
    })
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
root_dir <- knitr::opts_knit$get("root.dir")
if (!is.null(root_dir)){
    # This hack fixes the relative image paths. 
    # See https://github.com/rstudio/rmarkdown/issues/2473
    knitr::opts_knit$set(
        output.dir = root_dir
    )
}
proj_root <- rprojroot::find_package_root_file() |> normalizePath() 
# Utility function for figures to force them to have the correct path
find_figure <- function(names){
    rprojroot::find_package_root_file() |>
        file.path("man", "figures", names) 
}
```

<!-- badges: start -->

[![Lifecycle:experimental](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

**tidyprint** is an R package that provides a centralised tidy display strategy for biological data (e.g. SummarizedExperiment), and centralised messaging styles for the `tidyomics` packages. To facilitate the discussion about data display, we compare here four data diplay:

1.  **SummarizedExperiment**: Standard R/SummarizedExperiment printing style.

2.  **tidyprint_1**: Newly designed tibble abstraction, combines styles from **tidySummarizedExperiment** and **plyxp**

3.  **tidySummarizedExperiment**: Tidyverse-friendly tibble abstraction.

4.  **plyxp**: Top/bottom row truncation for large datasets.

Depending on your workflow and desired console output, `tidyprint` makes it easy to switch between these printing styles.

------------------------------------------------------------------------

## 1. Installation

You need the `remotes` package to install from GitHub. If you don’t have it, install via:

```{r eval=FALSE}
install.packages("remotes")
```

Then install **tidyprint** from GitHub:

```{r eval=FALSE}
remotes::install_github("tidyomics/tidyprint")
```

------------------------------------------------------------------------

## 2. Demo

Below is an example demonstrating how to use **tidyprint** with a sample `SummarizedExperiment` object.

### 2.1 Load Required Packages

```{r message = FALSE, warnings = FALSE}

library(dplyr)
library(tidyr)

# Now load tidyprint

library(tidyprint)

# Example SummarizedExperiment data from the airway package

data(se_airway)
se_airway

```

------------------------------------------------------------------------

### 2.2 **SummarizedExperiment** (Default)

By default, `print()` on a `SummarizedExperiment` displays the standard SummarizedExperiment info:

```{r}

# The default style:
se_airway %>% print()

```

```{r eval=FALSE}
# or equivalently:
se_airway %>% print(design = "SummarizedExperiment")
```

### 2.3 **tidyprint_1**

For a more compact view (top and bottom rows), similar to a plyxp/tidyverse style with tidySummarizedExperiment header and customised row_id:

```{r}

se_airway %>% print(design = "tidyprint_1")

```


### 2.4 **tidySummarizedExperiment**

Use the “tidySummarizedExperiment” design to view your data in a **tidy-friendly tibble** format:

```{r}

# Tidy SummarizedExperiment print:

se_airway %>% print(design = "tidySummarizedExperiment")

```

### 2.5 **plyxp**

For a more compact view (top and bottom rows), similar to a plyxp/tidyverse style:

```{r}

se_airway %>% print(design = "plyxp")

```

# Messaging function

We integrated a messaging function providing standardized, visually appealing messages for packages within the tidyomics ecosystem. It automatically detects the calling package to provide contextualized messaging, such as "tidyprint says" or "tidybulk says", enhancing consistency and readability across projects.

```{r}

test_tidy_message()

```


## Session info

```{r}

sessionInfo()

```
