---
title: "tidyprint"
output: "html_vignette"
params:
  demo_metadata: true
vignette: >
  %\VignetteIndexEntry{posteriorHCA}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
knit: >
    (function(x, ...){
        proj_root <- rprojroot::find_package_root_file() |> normalizePath()
        rmarkdown::render(
            x,
            output_format = "html_document",
            params = list(
                demo_metadata = TRUE
            )
        )
        rmarkdown::render(
            x,
            output_file = "README.md",
            output_format = "github_document",
            output_dir = proj_root,
            knit_root_dir = proj_root,
            params = list(
                demo_metadata = FALSE
            )
        )
    })
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
root_dir <- knitr::opts_knit$get("root.dir")
if (!is.null(root_dir)){
    # This hack fixes the relative image paths. 
    # See https://github.com/rstudio/rmarkdown/issues/2473
    knitr::opts_knit$set(
        output.dir = root_dir
    )
}
proj_root <- rprojroot::find_package_root_file() |> normalizePath() 
# Utility function for figures to force them to have the correct path
find_figure <- function(names){
    rprojroot::find_package_root_file() |>
        file.path("man", "figures", names) 
}
```

<!-- badges: start -->

[![Lifecycle:experimental](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

**tidyprint** is an R package that provides multiple printing styles for `SummarizedExperiment` objects. You can choose among:

1.  **SummarizedExperiment** (default): Standard R/SummarizedExperiment printing style.

2.  **tidySummarizedExperiment**: Tidyverse-friendly tibble abstraction.

3.  **plyxp**: Top/bottom row truncation for large datasets.

4.  **alternative_1**: Newly designed tibble abstraction, combines styles from **tidySummarizedExperiment** and **plyxp**

Depending on your workflow and desired console output, `tidyprint` makes it easy to switch between these printing styles.

------------------------------------------------------------------------

## 1. Installation

You need the \`\` package to install from GitHub. If you don’t have it, install via:

```{r eval=FALSE}
install.packages("devtools")
```

Then install **tidyprint** from GitHub:

```{r eval=FALSE}
devtools::install_github("tidyomics/tidyprint")
```

------------------------------------------------------------------------

## 2. Demo

Below is an example demonstrating how to use **tidyprint** with a sample `SummarizedExperiment` object.

### 2.1 Load Required Packages

```{r message = FALSE, warnings = FALSE}

library(dplyr)
library(tidyr)

# Now load tidyprint

library(tidyprint)

# Example SummarizedExperiment data from the airway package

data(se_airway)
se_airway

```

------------------------------------------------------------------------

### 2.2 **SummarizedExperiment** (Default)

By default, `print()` on a `SummarizedExperiment` displays the standard SummarizedExperiment info:

```{r}

# The default style:
se_airway %>% print()

```

```{r eval=FALSE}
# or equivalently:
se_airway %>% print(design = "SummarizedExperiment")
```

### 2.3 **tidySummarizedExperiment**

Use the “tidySummarizedExperiment” design to view your data in a **tidy-friendly tibble** format:

```{r}

# Tidy SummarizedExperiment print:

se_airway %>% print(design = "tidySummarizedExperiment")

```

### 2.4 **plyxp**

For a more compact view (top and bottom rows), similar to a plyxp/tidyverse style:

```{r}

se_airway %>% print(design = "plyxp")

```

You can also limit the number of displayed rows by setting `n_print` (or a similar argument in your code):

```{r}

se_airway %>% print(design = "plyxp", n_print = 5)

```

### 2.5 **alternative_1**

For a more compact view (top and bottom rows), similar to a plyxp/tidyverse style with tidySummarizedExperiment header and customised row_id:

```{r}

se_airway %>% print(design = "alternative_1")

```

You can also limit the number of displayed rows by setting `n_print` (or a similar argument in your code):

```{r}

se_airway %>% print(design = "alternative_1", n_print = 5)

```

## Session info

```{r}

sessionInfo()

```
